---
title: "Visualizing Sexism and Values"
subtitle: "STAT302: Data Visualization"
author: "Irene Kwon"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, comment = FALSE, dpi = 300, 
               echo = FALSE)

# Packages -----
library(cowplot)
library(ggrepel)
library(sf)
library(googleway)
library(ggspatial)
library(rnaturalearth)
library(rnaturalearthdata)
library(tidyverse)
library(rgeos)
library(skimr)
library(corrplot)

library(naniar)
library(mice)
library(finalfit)
library(GGally)
library(ggfittext)
## Read in the data -----
world <- ne_countries(scale = "medium", returnclass = "sf")
imputed_wvs <- read_rds("data/imputed_cart_comp.rds")

```

## Intro

Despite the contemporary waves of feminism and women's rights movements, sexism and gender-based oppression still persist in many parts of the world. For example, even in the modern days, more than one in four Africans see wife-beating as justifiable (Afrobarometer Survey: https://afrobarometer.org/press/more-one-four-africans-see-wife-beating-justifiable-afrobarometer-survey-finds), while more than 90% of the global population – including both men and women – is somehow biased against women (UNDP Gender Social Norms Index: https://www.undp.org/content/undp/en/home/news-centre/news/2020/Gender_Social_Norms_Index_2020.html). Yes, it is heartbreaking just to read about persistent misogyny - but to solve the problem, we have to begin by facing it directly. Therefore, for this project, I examine how sexist people around the world (on average) are, and whether tradition, religiosity, and religious fundamentalism are associated with one's sexism. 

Using the 6th wave of the WVS dataset, I prepared four sets of plots: (1) a histogram presenting the distribution of sexism, (2) three world maps colored to show how much sexist, conservative, and religious different countries are, (3) two boxplots to show the relationship between conservatism and sexism, and that between religiosity and sexism, and (4) a set of scatterplots to highlight the differences between the most conservative region and the most liberal region in terms of conservatism, religiosity, and fundamentalism. 

<br>

## Plot 1. Distribution of Sexism 

```{r plot1-histogram-sexism}
imputed_wvs %>%
  ggplot(aes(x = sexism)) + 
    geom_histogram(aes(y = ..density..), bins = 23) + 
    geom_vline(aes(xintercept = mean(sexism)), 
             color = "red", linetype = "dashed", size = 1) + 
    geom_density(alpha = .2, fill = "#1A5276") + 
    labs(
      title = "Distribution of Sexism", 
      caption = "Higher values indicate higher sexism.\n Red dotted line denotes the mean. \n(Minimum = 6, Maximum = 28, Mean = 14.23, Median = 14)", 
      subtitle = "World Value Survey, Wave 6"
  ) + 
    theme_minimal() + 
    theme(
      plot.title = element_text(size = 20, face = "bold", hjust = 0.5), 
      plot.subtitle = element_text(hjust = 0.5)
  ) 

```


## Plot 2. World Maps: Sexism, Conservatism, and Religiosity around the World 

```{r plot1b-sexism-worldmap}
# data prep for plot 1 (to combine with the world dataset)
wvs2 <- imputed_wvs %>%
  mutate(sexism_imputed = sexism, 
         sexism = men_more_right_jobs + men_better_poli_leaders + men_better_biz_exec + 
           uni_edu_boys + women_making_more_problem + dem_women_same_right_men)

wvs_comb <- wvs2 %>%
  group_by(countryname) %>% 
  summarize(sexism_mean = mean(sexism)) %>% arrange(desc(sexism_mean)) %>% 
  rename(subunit = countryname) # 60 country names 

tibble <- left_join(world, wvs_comb, by = "subunit") %>% 
  as_tibble() %>% select(subunit, sexism_mean) %>% 
  na.omit() %>% 
  arrange(sexism_mean) # 55 observations 

## Look at how the "world" dataset writes their name: 
#worldtibble <- world %>% as_tibble() 
#worldtibble$subunit %>% table()
##### country names in the 'world' dataset: South Korea, Hong Kong S.A.R., Russia, Palestine, Taiwan 

## substitute the value with that name in the wvs_comb data file 
wvs_comb[32, 1] <- "Russia"
wvs_comb[34, 1] <- "South Korea"
wvs_comb[2, 1] <- "Palestine"
wvs_comb[39, 1] <- "Taiwan"
wvs_comb[36, 1] <- "Hong Kong"

## then left_join the re-coded one 
tibble <- left_join(world, wvs_comb, by = "subunit") %>% 
  as_tibble() %>% select(subunit, sexism_mean) %>% 
  na.omit() %>% 
  arrange(sexism_mean)

## then try coloring the map again 
combined_world <- left_join(world, wvs_comb, by = "subunit") %>% 
  st_as_sf()

## Exclude Antarctica from the map
combined_world2 <- combined_world[combined_world$subunit != "Antarctica", ]

## Plot 1 sexism 
ggplot(data = combined_world2, aes(geometry = geometry)) + 
  geom_sf(aes(fill = sexism_mean)) + 
    theme_map() + 
    scale_fill_gradient(high = "darkred",
                      low = "lightpink", 
                      name = "average sexism", 
                      # setting breaks gets rid of the legend - this looks nicer!
                      breaks = c(9, 18)) + 
    labs(
      title = "Sexism Around the World", 
      subtitle = "World Values Survey, 2010-2014", 
      caption = "Higher values of average sexism indicate greater sexism. \n The brighter a country is colored, the less sexist its peaple are on average. \n Gray areas indicate no data available for the given country."
  ) + 
    theme(
      plot.title = element_text(hjust = 0.5), 
      plot.subtitle = element_text(hjust = 0.5)
  )
```


```{r data-wrangling}
conservatism <- imputed_wvs %>% 
  select(countryname, continent, sub_region, sex, 
         sexism, just_sexbeforemarriage, just_homosexuality,
         parentsproudlifegoal, 
         neighbor_homosexual, neighbor_unmarriedcouple, housewifefulfilling, 
         adventurerisk, behaveproperlyimpt, traditionimportant,
         abortionjustifiable, 
         workingmomchildsuffer, respectforelder, respectforauthority,
         divorcejustifiable, 
         obedienceimportant, boss30yrold)

# re-code the variables so that lower values mean more conservatism 
conservatism <- conservatism %>%
  mutate(neighbor_homosexual = 1 - neighbor_homosexual, 
         neighbor_unmarriedcouple = 1 - neighbor_unmarriedcouple, 
         adventurerisk = 7 - adventurerisk, 
         obedienceimportant = 1 - obedienceimportant)

# create the combined index score for conservatism 
conservatism2 <- conservatism %>% 
  mutate(conservatism = just_sexbeforemarriage + just_homosexuality +
           parentsproudlifegoal + 
           neighbor_homosexual + neighbor_unmarriedcouple + 
           housewifefulfilling + 
           adventurerisk + behaveproperlyimpt + traditionimportant +
           abortionjustifiable + 
           workingmomchildsuffer + respectforelder + respectforauthority +
           divorcejustifiable + 
           obedienceimportant + boss30yrold) %>% 
  select(countryname, continent, sub_region, sex, sexism, conservatism)

# create a factor variable for conservatism 
conservatism2 <- conservatism2 %>% 
  mutate(conservatism_cat = ifelse(conservatism < 25, 1,
                                   ifelse(conservatism < 35, 2, 
                                          ifelse(conservatism < 45, 3, 
                                                 ifelse(conservatism < 55, 4, 
                                                        ifelse(conservatism < 65, 5, 
                                                               ifelse(conservatism < 75, 6, 7)))))))

conservatism2 <- conservatism2 %>% 
  mutate(conservatism_cat = as_factor(conservatism_cat))

# Re-code the variables so that the lower values mean the more religiosity
imputed_wvs <- imputed_wvs %>% 
  mutate(
    # in a 1, 2 binary variable (1 mentioned, 2 not mentioned)
    religion_childquality2 = 2 - religion_childquality, 
    # into a 1-3 variable: 1 active member, 3 not a member
    religiousmembership2 = 3 - religiousmembership, 
    sciencevsreligion2 = 5 - sciencevsreligion, 
    toomuchsciencelessfaith2 = 11 - toomuchsciencelessfaith, 
    believeinhell2 = 2 - believeinhell, 
    godimportant2 = 11 - godimportant) %>% 
  mutate(religiosity = religion_important + 
           religion_childquality2 + religiousperson + 
           attendreligion + oftenpray + religiousmembership2 + 
           sciencevsreligion2 + 
           toomuchsciencelessfaith2 + believeinhell2 + godimportant2)

# Re-code the variables so that the lower values mean the more religious fundamentalism
imputed_wvs <- imputed_wvs %>% 
  mutate(teachallreligion2 = 5 - teachallreligion, 
         differentreligionmoral2 = 5 - differentreligionmoral, 
         neighbor_diffreligion2 = 2 - neighbor_diffreligion, 
         trust_diffreligion2 = 5 - trust_diffreligion) %>% 
  mutate(
    fundamentalism = onlyacceptablereligion + teachallreligion2 +
      differentreligionmoral2 + 
      neighbor_diffreligion2 + trust_diffreligion2
  )

# combine to construct a dataset 
religion <- imputed_wvs %>% select(fundamentalism, religiosity)
selected_dat <- cbind(conservatism2, religion) %>% as_tibble()

```


```{r plot1c-conservatism-worldmap}
## CREATE THE MAP 
comb_conservatism <- selected_dat %>%
  group_by(countryname) %>% 
  summarize(conservatism_mean = mean(conservatism)) %>% 
  arrange(desc(conservatism_mean)) %>% 
  rename(subunit = countryname) 

comb_conservatism[17, 1] <- "Russia"
comb_conservatism[11, 1] <- "South Korea"
comb_conservatism[50, 1] <- "Palestine"
comb_conservatism[19, 1] <- "Taiwan"
comb_conservatism[20, 1] <- "Hong Kong S.A.R."

combined_world_conservatism <- left_join(world, comb_conservatism, by = "subunit") %>% 
  st_as_sf()

## Exclude Antarctica from the map
combined_world_conservatism2 <- combined_world_conservatism[combined_world_conservatism$subunit != "Antarctica", ]

## Plot 1. World Map 
ggplot(data = combined_world_conservatism2, 
                             aes(geometry = geometry)) + 
  geom_sf(aes(fill = conservatism_mean)) + 
  scale_fill_gradient(low = "darkblue",
                      high = "lightskyblue1", 
                      name = "average sexism", 
                      # setting breaks gets rid of the legend - this looks nicer!
                      breaks = c(9, 18)) + 
  labs(
    title = "Conservatism Around the World", 
    subtitle = "World Values Survey, 2010-2014", 
    caption = "The brighter a country is colored, the less conservative its peaple are on average. \n Gray areas indicate no data available for the given country."
  ) + 
  theme_map() + 
  theme(
    plot.title = element_text(hjust = 0.5), 
    plot.subtitle = element_text(hjust = 0.5)
  )

```


```{r plot1d-world-map-religiosity}
## Data wrangling for creating the world map
comb_religiosity <- selected_dat %>%
  group_by(countryname) %>% 
  summarize(religiosity_mean = mean(religiosity)) %>% 
  arrange(desc(religiosity_mean)) %>% 
  rename(subunit = countryname) 

comb_religiosity[12, 1] <- "Russia"
comb_religiosity[15, 1] <- "South Korea"
comb_religiosity[44, 1] <- "Palestine"
comb_religiosity[17, 1] <- "Taiwan"
comb_religiosity[6, 1] <- "Hong Kong S.A.R."


## World Map of Religiosity
combined_world_religiosity<- left_join(world, comb_religiosity, 
                                       by = "subunit") %>% 
  st_as_sf()

combined_world_religiosity2 <- combined_world_religiosity[combined_world_religiosity$subunit != "Antarctica", ]

ggplot(data = combined_world_religiosity2, 
                            aes(geometry = geometry)) + 
  geom_sf(aes(fill = religiosity_mean)) + 
  scale_fill_gradient(low = "#286101",
                      high = "#adde8c", 
                      name = "average religiosity", 
                      # setting breaks gets rid of the legend - this looks nicer!
                      breaks = c(9, 18)) + 
  labs(
    title = "Religiosity Around the World", 
    subtitle = "World Values Survey, 2010-2014", 
    caption = "The brighter a country is colored, the less religious its peaple are on average. \n Gray areas indicate no data available for the given country."
  ) + 
  theme_map() + 
  theme(
    plot.title = element_text(hjust = 0.5), 
    plot.subtitle = element_text(hjust = 0.5)
  )

```


<br>

## Plot 3. Relational Distribution: Sexism, Conservatism and Religiosity

```{r plot2-data-wrangling}
# labeling dataset 
# (1) male
class_dat_male <- conservatism2 %>%
  filter(sex == 1) %>% 
  group_by(conservatism_cat) %>% 
  summarise(
    n = n(), 
    sexism = mean(sexism), 
    label = str_c("n = ", n, sep = "")
  )

# (2) female 
class_dat_female <- conservatism2 %>%
  filter(sex == 2) %>% 
  group_by(conservatism_cat) %>% 
  summarise(
    n = n(), 
    sexism = mean(sexism), 
    label = str_c("n = ", n, sep = "")
  )

# (3) entire sex 
class_dat <- conservatism2 %>%
  group_by(conservatism_cat) %>% 
  summarise(
    n = n(), 
    sexism = mean(sexism), 
    label = str_c("n = ", n, sep = "")
  )

```


```{r plot2a-boxplot-sexism-conservatism}
conservatism2 %>%
  ggplot(aes(x = conservatism_cat, y = sexism)) + 
    geom_boxplot(aes(group = conservatism_cat), varwidth = TRUE) + 
  # add mean sexism point for female
    geom_point(data = class_dat_female, 
             aes(x = conservatism_cat, y = sexism), 
             color = "red", alpha = 0.5, 
             size = 5) + 
  # add mean sexism point for male 
    geom_point(data = class_dat_male, 
             aes(x = conservatism_cat, y = sexism), 
             color = "darkblue", alpha = 0.5, size = 5) + 
    geom_text(data = class_dat, aes(label = label, y = 2)) + 
    scale_x_discrete(labels = c("most conservative", "6", "5", "4", "3", "2", "least conservative")) + 
    theme_minimal() + 
    labs(
      title = "Conservatism and Sexism",
      x = "Conservatism", 
      y = "Sexism", 
      caption = "Red dots denote the mean sexism score for women in the given category. \n Blue dots denote the mean sexism score for men in the given category."
  ) + 
    theme(
      plot.title = element_text(hjust = 0.5, size = 20, face = "bold"), 
      axis.title = element_text(face = "bold"), 
      axis.text = element_text(size = 12)
  ) 

```



```{r plot2b-boxplot-sexism-religiosity}
# create a factor variable for religiosity 
selected_dat2 <- selected_dat %>% 
  mutate(religiosity_cat = ifelse(religiosity < 15, 1,
                                   ifelse(religiosity < 20, 2, 
                                          ifelse(religiosity < 25, 3, 
                                                 ifelse(religiosity < 30, 4, 
                                                        ifelse(religiosity < 35, 5, 
                                                              ifelse(religiosity < 40, 6, 
                                                                     ifelse(religiosity < 45, 7, 
                                                                            ifelse(religiosity < 50, 8, 9)))))))))

selected_dat2 <- selected_dat2 %>% 
  mutate(religiosity_cat = as_factor(religiosity_cat))


## label info 
# (1) male
class_dat_male_religiosity <- selected_dat2 %>%
  filter(sex == 1) %>% 
  group_by(religiosity_cat) %>% 
  summarise(
    n = n(), 
    sexism = mean(sexism), 
    label = str_c("n = ", n, sep = "")
  )

# (2) female 
class_dat_female_religiosity <- selected_dat2 %>%
  filter(sex == 2) %>% 
  group_by(religiosity_cat) %>% 
  summarise(
    n = n(), 
    sexism = mean(sexism), 
    label = str_c("n = ", n, sep = "")
  )

# (3) entire sex 
class_dat_religiosity <- selected_dat2 %>%
  group_by(religiosity_cat) %>% 
  summarise(
    n = n(), 
    sexism = mean(sexism), 
    label = str_c("n = ", n, sep = "")
  )

```



```{r plot3-plotting}
## PLOT 
selected_dat2 %>%
  ggplot(aes(x = religiosity_cat, y = sexism)) + 
    geom_boxplot(aes(group = religiosity_cat), varwidth = TRUE) + 
  # add mean sexism point for female
    geom_point(data = class_dat_female_religiosity, 
             aes(x = religiosity_cat, y = sexism), 
             color = "red", alpha = 0.5, 
             size = 5) + 
  # add mean sexism point for male 
    geom_point(data = class_dat_male_religiosity, 
             aes(x = religiosity_cat, y = sexism), 
             color = "darkblue", alpha = 0.5, size = 5) + 
    geom_text(data = class_dat_religiosity, aes(label = label, y = 2)) + 
    scale_x_discrete(labels = c("most religious", "8", "7", "6", "5", "4", 
                                "3", "2", "least religious")) + 
    labs(
      title = "Religiosity and Sexism",
      x = "Religiosity", 
      y = "Sexism", 
      caption = "Red dots denote the mean sexism score for women in the given category. \n Blue dots denote the mean sexism score for men in the given category."
  ) + 
    theme_minimal() + 
    theme(
      plot.title = element_text(hjust = 0.5, size = 20, face = "bold"), 
      axis.title = element_text(face = "bold"), 
      axis.text = element_text(size = 12)
      ) 

```

<br>


## Plot 4. Scatterplots

```{r plot4-scatterplot-data-wrangle}
south_asia <- selected_dat %>% 
  filter(sub_region == "Southern Asia")
northern_europe <- selected_dat %>% 
  filter(sub_region == "Northern Europe")
western_europe <- selected_dat %>% 
  filter(sub_region == "Western Europe")
western_africa <- selected_dat %>% 
  filter(sub_region == "Western Africa")
northern_africa <- selected_dat %>% 
  filter(sub_region == "Northern Africa")
```




```{r plot4a-creating-the-scatterplot}
## MOST CONSERVAtIVE: SOUTHERN ASIA, MOST LIBERAL: WESTERN EUROPE 
a <- sample_frac(selected_dat, size = 0.3, replace = FALSE) %>% 
  ggplot(aes(conservatism, sexism)) + 
  geom_point(color = "grey") +
  geom_jitter(color = "grey") + 
  geom_point(data = sample_frac(south_asia, size = 0.3, replace = FALSE), 
             aes(conservatism, sexism), color = "red", alpha = 0.6) +
  geom_jitter(data = sample_frac(south_asia, size = 0.3, replace = FALSE), 
              aes(conservatism, sexism), color = "red", alpha = 0.6) + 
  geom_point(data = sample_frac(western_europe, size = 0.3, replace = FALSE), 
             aes(conservatism, sexism), color = "darkblue", alpha = 0.6) + 
  geom_jitter(data = sample_frac(western_europe, size = 0.3, replace = FALSE), 
             aes(conservatism, sexism), color = "darkblue", alpha = 0.6) + 
  theme_minimal() + 
  theme(
    axis.title.x = element_text(face = "bold"), 
    axis.title.y = element_text(face = "bold")
  )

a <- ggdraw(add_sub(plot = a, 
                 label = "South Asia (red)\nWestern Europe (blue)", 
                 size = 10))
```


```{r plot4b-creating-the-scatterplot}
## MOST RELIGIOUS: WESTERN AFRICA, LEAST RELIGIOUS: NORTHERN EUROPE
b <- sample_frac(selected_dat, size = 0.3, replace = FALSE) %>% 
  ggplot(aes(religiosity, sexism)) + 
  geom_point(color = "grey") +
  geom_jitter(color = "grey") + 
  geom_point(data = sample_frac(western_africa, size = 0.3, replace = FALSE), 
             aes(religiosity, sexism), color = "red", alpha = 0.6) +
  geom_jitter(data = sample_frac(western_africa, size = 0.3, replace = FALSE), 
              aes(religiosity, sexism), color = "red", alpha = 0.6) + 
  geom_point(data = sample_frac(northern_europe, size = 0.3, replace = FALSE), 
             aes(religiosity, sexism), color = "darkblue", alpha = 0.6) + 
  geom_jitter(data = sample_frac(northern_europe, size = 0.3, replace = FALSE), 
              aes(religiosity, sexism), color = "darkblue", alpha = 0.6) + 
  theme_minimal() + 
  theme(
    axis.title.x = element_text(face = "bold"), 
    axis.title.y = element_text(face = "bold")
  )

b <- ggdraw(add_sub(plot = b, 
                 label = "Western Africa (red)\nNorthern Europe (blue)", 
                 size = 10))
```



```{r plot4c-creating-the-scatterplot}
## FUNDAMENTALISM
c <- sample_frac(selected_dat, size = 0.3, replace = FALSE) %>% 
  ggplot(aes(fundamentalism, sexism)) + 
  geom_point(color = "grey") +
  geom_jitter(color = "grey") + 
  geom_point(data = sample_frac(northern_africa, size = 0.3, replace = FALSE), 
             aes(fundamentalism, sexism), color = "red", alpha = 0.6) +
  geom_jitter(data = sample_frac(northern_africa, size = 0.3, replace = FALSE), 
              aes(fundamentalism, sexism), color = "red", alpha = 0.6) + 
  geom_point(data = sample_frac(northern_europe, size = 0.3, replace = FALSE), 
             aes(fundamentalism, sexism), color = "darkblue", alpha = 0.6) + 
  geom_jitter(data = sample_frac(northern_europe, size = 0.3, replace = FALSE), 
              aes(fundamentalism, sexism), color = "darkblue", alpha = 0.6) + 
  theme_minimal() + 
  theme(
    axis.title.x = element_text(face = "bold"), 
    axis.title.y = element_text(face = "bold")
  )

c <- ggdraw(add_sub(plot = c, 
                 label = "Northern Africa (red)\nNorthern Europe (blue)", 
                 size = 10))
```



```{r plot4-printing-out-the-plot}
plot4 <- cowplot::plot_grid(a, b, c, nrow = 1)

# now add the title
title <- ggdraw() + 
  draw_label(
    "The Most and the Least Sexist Regions in the World\n in Relation to Traditional Values",
    fontface = 'bold',
    x = 0,
    hjust = 0 
  ) + 
  theme(
    # to arrange the title 
    plot.margin = margin(0, 0, 0, 10)
  )

plot4 <- plot_grid(
  title, plot4,
  ncol = 1,
  # rel_heights values control vertical title margins
  rel_heights = c(0.1, 1)
)

ggdraw(add_sub(plot4, 
               "Each scatterplot contrasts the most (colored in red) and the least (colored in blue)\nconservative, religious, and religiously fundamental regions, respectively.", 
               size = 10))
```


## Conclusion 

The previous plots suggest a couple of expected, yet interesting findings. First, the histogram showing the distribution of sexism around the world reveals that although most people around the world are "moderately" sexist (denoted by the fat middle), some people are extremely sexist (observed from the long right tail). 

Second, we do observe the regional variations. From the three world maps presented above, we can see that on average, people in the Americas (including northern, central and Latin), Oceania, and European (mainly Western and Northern) countries are less sexist. People in African and Middle Eastern countries are, on the other hand, the most sexist on average. Asians scored about the middle. We can see a similar pattern in the distribution of conservatism around the world. Middle Eastern and African countries were more traditional/conservative than other parts of the world while western and northern European countries are notably less so. Given that similar pattern is observed in the sexism map, it does confirm our priors that more traditional/conservative societies are more likely to be sexist. Probably this may be because tradition defines gender roles more rigidly, and therefore, people socialized in traditional societies are more likely to prescribe certain norms to women (or to men) more rigidly. 

Since religion has been around probably ever since the beginning of human race - be it Christianity, Muslim, Hinduism, Buddhism, or less institutionalized tribal religions - I was curious whether religion has something to do with sexist beliefs. While they believe in different deity, a lot of religions actually share sexism. We also see the similar pattern for the religiosity map - the more conservative a given society is, the more religious it is. An exception is China, where religion is nationally band (hence less religious). And these coloring patterns are also associated with the average extent of the societal sexism. We can see the similar pattern at the last set of scatterplots. These scatterplots are meant to highlight differnet societies (i.e., regions) at the extreme ends of the X-axis variables (conservatism, religiosity, and religious fundamentalism). Interestingly, the more conservative, religious, and fundamental a society is, the more sexist it is as well. 

The boxplots show that women are, on average, less sexist than men are - net of religiosity and conservatism. The boxplots also reveal the association between sexism, conservatism, and religiosity. 

While this is a visualization-oriented exploratory analysis, and by no means do I intend to reach a conclusive statement about the relationship between religion, conservatism, tradition and sexism. Still, I believe the data visualization in this report can be a good starting point of more nuanced understanding of our values. 


